---
title: "Assignment 1 for Data Analysis 2 and Coding with R"
date: 2020-11-24
author: "Pauline Broussolle"
output: 
  pdf_document:
  toc: true
  toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load packages
library(tidyverse)
require(scales)
library(lspline)
library(estimatr)
library(texreg)
library(ggthemes)
library(WDI)
library(magrittr)
library(moments)
library(gridExtra)

# Import data
data_in <- "~/Documents/GitHub/Assignment-1/data/"
df <- read_csv(paste0(data_in,"clean/covid_pop_10_15_2020_clean.csv"))
```

## Introduction

The aim of the analysis is to analyse the pattern of association between registered covid-19 cases and registered number of death due to covid-19 on a **15/10/2020**. 
My dependent variable (y) is the **number of registered death due to covid** and my explanatory variable (x) is the **number of registered cases**. The aim of the analysis is to create a report on the pattern of association, choose and interpret a regression model and refer to robustness checks.

Introduction of the dataset and variables
The aim of this assignment is to guide you through in creating a short report on the pattern of association. You will need to chose your final model, interpret the results and  
### Executive summary

## 1. Summary statistics and Distribution for x and y

```{r , echo = FALSE }
confirmed_cases <- df %>% 
  summarise(
  mean     = round( mean(confirmed), 0),
  median   = median(confirmed),
  min      = min(confirmed),
  max      = max(confirmed),
  std      = round( sd(confirmed), 0))

nb_of_death <- df %>% 
  summarise(
  mean     = round( mean(death), 0),
  median   = median(death),
  min      = min(death),
  max      = max(death),
  std      = round( sd(death), 0))


summary <- confirmed_cases %>% add_row( nb_of_death)
summary <- mutate (summary, variable = c("confirmed cases" , "nb of death"))
knitr::kable(summary,caption="Summary for the number of registered death and registered cases")
```

2-3 sentence, explain the main features and distribution - use histograms and summary statistics table (mean, median, min, max, standard deviation)
 
```{r , echo = FALSE, out.width = '50%' }
# Histograms
ggplot(data = df, aes(x=death)) + 
  geom_histogram(binwidth =1000,col="black", alpha=0.7)+
  labs(x="number of registered death due to covid", y="Absolute Frequency") +
  theme_bw()

ggplot(data = df, aes(x=confirmed)) + 
  geom_histogram(binwidth =10000, fill= "darkorange",col="black", alpha=0.7)+
  labs(x="number of registered cases", y="Absolute Frequency") +
  theme_bw()
```

### Select or drop observations, checking extreme values
```{r , include = FALSE }
df %>% filter( confirmed > 2000000, death > 50000 )
```
We check countries which have confirmed cases aboves 2 milion and registered number of death above 50,000. These are India, Brazil and United States, which are not measurement errors. We keep these values.

## 2. Investigate the transformation of your variables
### Scaling

```{r , echo = FALSE, out.width = '33%'  }
# Check basic scatter-plots!
#   Two competing models:
#     1) death = alpha + beta * confirmed
#     2) death = alpha + beta * confirmedsc
#
# Where to use log-transformation? - level-level vs level-log vs log-level vs log-log
#
# 1) lifeexp - gdptot: level-level model without scaling
ggplot( df , aes(x = confirmedsc, y = death)) +
  geom_point() +
  geom_smooth(method="loess")+
  labs(x = "nb of confirmed cases",y = "nb of death") 

# The pattern seems non-linear and there are some extreme observations, corresponding to China, US, India

# You can change the scale for Total GDP for checking log-transformation
ggplot( df , aes(x = confirmed, y = death)) +
  geom_point() +
  geom_smooth(method="loess")+
  labs(x = "nb of confirmed cases",y = "nb of death")+
  scale_x_continuous( trans = log_trans(),  breaks = c(1,2,5,10,20,50,100,200,500,1000,10000) )

# You can change the scale for Total GDP and life-expectancy for checking log-log transformation
ggplot( df , aes(x = confirmed, y = death)) +
  geom_point() +
  geom_smooth(method="loess")+
  labs(x = "nb of confirmed cases",y = "nb of death")+
  scale_x_continuous( trans = log_trans(),  breaks = c(1,2,5,10,20,50,100,200,500,1000,10000) )+
  scale_y_continuous( trans = log_trans() )
```

Estimating different models

Presentation of model choice 

Hypothesis testing on beta (which interacts with x)

Analysis of the residuals

Conclusion

