---
title: "Assignment 1 for Data Analysis 2 and Coding with R"
date: 2020-11-24
author: "Pauline Broussolle"
output: 
  pdf_document:
  toc: true
  toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load packages
library(tidyverse)
require(scales)
library(lspline)
library(estimatr)
library(texreg)
library(ggthemes)
library(WDI)
library(magrittr)
library(moments)
library(gridExtra)

# Import data
data_in <- "~/Documents/GitHub/Assignment-1/data/"
df <- read_csv(paste0(data_in,"clean/covid_pop_10_15_2020_clean.csv"))
```

## Introduction

My goal is to analyse the pattern of association between registered covid-19 cases and registered number of death due to covid-19 on **15/10/2020**. 
My dependent variable (y) is the **number of registered death due to covid** and my explanatory variable (x) is the **number of registered cases**. The aim of the analysis is to create a report on the pattern of association, choose and interpret a regression model and refer to robustness checks.

Introduction of the dataset and variables
The aim of this assignment is to guide you through in creating a short report on the pattern of association. You will need to chose your final model, interpret the results and  
## Executive summary

## 1. Preparatory Data Analysis

### Summary statistics and Distribution for x and y

```{r , include = FALSE }
summary(df)
```


```{r , echo = FALSE }
confirmed_cases <- df %>% 
  summarise(
  mean     = round( mean(confirmed), 0),
  median   = median(confirmed),
  min      = min(confirmed),
  max      = max(confirmed),
  std      = round( sd(confirmed), 0))

nb_of_death <- df %>% 
  summarise(
  mean     = round( mean(death), 0),
  median   = median(death),
  min      = min(death),
  max      = max(death),
  std      = round( sd(death), 0))


summary <- confirmed_cases %>% add_row( nb_of_death)
summary <- mutate (summary, variable = c("confirmed cases" , "nb of death"))
knitr::kable(summary,caption="Summary for the number of registered death caused by covid and registered covid cases")
```

2-3 sentence, explain the main features and distribution - use histograms and summary statistics table (mean, median, min, max, standard deviation)
 
```{r , echo = FALSE, out.width = '50%' }
# Histograms
ggplot(data = df, aes(x=death)) + 
  geom_histogram(binwidth =1000,col="black", alpha=0.7)+
  labs(x="number of registered death due to covid", y="Absolute Frequency") +
  theme_bw()

ggplot(data = df, aes(x=confirmed)) + 
  geom_histogram(binwidth =10000, fill= "darkorange",col="black", alpha=0.7)+
  labs(x="number of registered covid cases", y="Absolute Frequency") +
  theme_bw()
```

### Select or drop observations, checking extreme values
```{r , include = FALSE }
df %>% filter( confirmed > 2000000, death > 50000 )
```
We check countries which have confirmed cases aboves 2 milion and registered number of death above 50,000. These are India, Brazil and United States, which are not measurement errors. We keep these values.

## 2. Investigate the transformation of the variables
### Scaling, Take Logs?

```{r , echo = FALSE, fig.height=3.0, warning = FALSE, message = FALSE}
# The model investigated is: death = alpha + beta * confirmed

# Check the possible different ln transformation for the variables with plotting different scatter- plots with lo(w)ess.

# 1) Nb of death - confirmed cases: level-level model without scaling
p1 <- ggplot( df , aes(x = confirmed, y = death)) +
  geom_point() +
  geom_smooth(method="loess")+
  labs(x = "confirmed cases",y = "nb of death") 

# Change the scale for confirmed cases for checking log-transformation
p2 <- ggplot( df , aes(x = confirmed, y = death)) +
  geom_point() +
  geom_smooth(method="loess")+
  labs(x = "confirmed cases,ln scale",y = "nb of death")+
  scale_x_continuous( trans = log_trans(),  breaks = c(1,2,5,10,20,50,100,200,500,1000,10000) )

# Change the scale for confirmed cases and nb of death for checking log-log transformation
p3 <- ggplot( df , aes(x = confirmed, y = death)) +
  geom_point() +
  geom_smooth(method="loess")+
  labs(x = "confirmed cases,ln scale",y = "nb of death, ln scale")+
  scale_x_continuous( trans = log_trans(),  breaks = c(1,2,5,10,20,50,100,200,500,1000,10000) )+
  scale_y_continuous( trans = log_trans() )

library(gridExtra)
grid.arrange(p1, p2, p3 , ncol=3, nrow = 1)

```

For the simple model without scaling the pattern is non-linear, most of observations are concentrated and there are some extreme observations, corresponding to Brazil, US, India.

Make a substantive and statistical reasoning, where and when to use ln transformation. You do not need to fit any model here, only use statistical reasoning based on the graphs.
i. Take care when it is possible to make ln transformation: you may need to drop or change some variables.

 2) using only gdppc is possible, but need to model the non-linearity in data
     - Substantive: Level changes is harder to interpret and our aim is not to get $ based comparison
     - Statistical: log transformation is way better approximation make simplification!
  3) taking log of confirmed cases and log of death is making the association close to linear!
 4) taking log for life-expectancy does not matter -> use levels!
    - Substantive: it does not give better interpretation
    - Statistical: you can compare models with the same y, no better fit
    - Remember: simplest the better!

We should use the log-log transformation, which is the only to provide a linear pattern.

We create new variables which are ln_confirmed and ln_death .
```{r , include = FALSE}
# Take Log of gdp/capita and log GDP total
df <- df %>% mutate( ln_confirmed = log( confirmed ),
                     ln_death = log( death ) )
```

## 3. Estimating different models
I chose the log-log transformation : ln_death = alpha + beta*ln_confirmed

### Simple linear regression
```{r , echo = FALSE }
reg1 <- lm_robust (death ~ ln_confirmed  , data = df, se_type = "HC2" )
summary( reg1 )
ggplot( data = df, aes( x = ln_confirmed, y = ln_death ) ) + 
  geom_point( color='blue') +
  geom_smooth( formula = y ~ poly(x,2) , method = lm , color = 'red' )
```

### Quadratic regression
```{r , include = FALSE, out.width = '50%' }
# Take Log of gdp/capita and log GDP total
df <- df %>% mutate( ln_confirmed = log( confirmed ),
                     ln_death = log( death ) )
```

### Piecewise linear spline regression
```{r , include = FALSE, out.width = '50%' }
# Take Log of gdp/capita and log GDP total
df <- df %>% mutate( ln_confirmed = log( confirmed ),
                     ln_death = log( death ) )
```

### Weighted linear regression, using population as weights. 
```{r , include = FALSE, out.width = '50%' }
# Take Log of gdp/capita and log GDP total
df <- df %>% mutate( ln_confirmed = log( confirmed ),
                     ln_death = log( death ) )
```


### Presentation of model choice 
(i) Compare the models and choose your preferred one

Hypothesis testing on beta (which interacts with x)

Analysis of the residuals

Conclusion

